Covering Your Basis: Model Agnostic Probes of the Binary Black Hole Population

### 2022-09-26 src/tex/ms.tex
move macros
update readme
add abstract

\author{Jaxen Godfrey}
\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}

\author{Alexis Vives}
\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}

\author{Jon Hornung}
\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}
\affiliation{Thurston High School, Springfield, OR 9747, USA}

\begin{abstract}
We introduce a non-parametric model built from basis splines, and apply it to infer the astrophysical distribution of 
binary black hole masses and spins using the 69 confident observations in GWTC-3. We find the same features previously recovered 
with similarly flexible models in the mass distribution, namely the peaks in merger rates at \~10\msun and \~35\msun. 
We apply our method to investigate whether the spin components are indendendenty distributed with a model agnositc way, where we find that 
the spin magnitude distribution seems to be quite consistently IID, with XX\% of black hole spins below XX at XX\% credibility. We report 
slight evidence of tilt distributions that do not peak at $\cos(\theta_i) = 1$ as previous models assumed, and slight evidence that 
the components are distributed differently, in contrast to the spin magnitude distribution. While we do not find signs of a significant
zero spinning sub population we do find that the data prefer not to force the support for exactly zero spinning black holes to 0 as 
the non-singular Beta distribution and \textsc{Default} model assumes. \result{$\CIPlusMinus{\macros[MSPlineIIDCompSpins][peakCosTilt]}$}
\result{$\CIPlusMinus{\macros[MSPlineIndependentCompSpins][peakCosTilt1]}$}, \result{$\CIPlusMinus{\macros[MSPlineIndependentCompSpins][peakCosTilt2]}$}
\end{abstract}

>>>>

%\author{Jaxen Godfrey}
%\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}

%\author{Alexis Vives}
%\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}

%\author{Jon Hornung}
%\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}
%\affiliation{Thurston High School, Springfield, OR 9747, USA}

\begin{abstract}                 
We introduce a non-parametric model built from basis splines, and apply it to infer the astrophysical distribution of 
binary black hole (BBH) masses and spins using the 69 confident observations in GWTC-3. We report the same features previously recovered 
with similarly flexible models in the mass distribution, most notably the peaks in merger rates at \~10\msun and \~35\msun. 
We then use our method as the first non-parametric model applied to the BBH spin distribution, corroborating the results of \citet{spinitasyoulike}, 
showing that tilt distributions may not peak at $\cos(\theta_i) = 1$ as previous models assumed. We find broad agreement with the 
accepted conlcusions of the spin magnitude distribution: the majority of BBH spins are small ($a<0.5$), the distribution peaks at $a\sim0.2$, 
and there is mild support for a non-spinning subpopulaton, but cannot resolve this with current catalog sizes. We also test components whether the data 
prefers the spin distributions to be identical for each component, finding no strong sign of differening distibutions. When modeling the $\chi_\mathrm{eff}$ 
distribution with basis splines we find similar conclusions as previously reported, consistent with a non-isotropic distribution 
(peaks at $\chi_\mathrm{eff}$\result{$\CIPlusMinus{\macros[ChiEffective][chieff][PeakChiEff]}$}), showing considerable support 
($\chi_\mathrm{eff}=$\result{$\CIPlusMinus{\macros[ChiEffective][chieff][FracBelow0]}$}) for negative effective spins or misaligned systems.
We conclude by discussing how non-parametric methods in gravitatonal-wave population inference are poised to provide a useful complement to 
the usual parametric approach going into the data rich catalog era of gravitational-wave astronomy.
\end{abstract}
